Baseline centering problem

### 2018-12-13 paper/tangotd.tex

``Universal value function approximators'' (Schaul--Horgan--Gregor--Silver ICML 2015) do a bit of that, but only
with the left (ordinary) Bellman equation. It amounts to solving $\#S$
independent Bellman equations, and relying on the generalization
abilities of the network, but the right equation is not exploited at all.

@@

If the state space is finite and if we use a one-hot representation, a
solution is indeed $E=\R^{\#S}$ with
\begin{equation}
f(s_2)=\1_{s_2},\qquad
(\green(s_1))_{s_2}=(\Delta^{-1})_{s_1s_2}/\rho(s_2)
\end{equation}

The reason for this form is the following: $E$ is supposed to represent
distributions/functions over states. $f(s_2)$ is supposed to be
mapped to $\1_{s_2}$. $\green(s_1)$ is supposed to be mapped to the
density wrt $\rho$ of the fundamental matrix at $s_1$ (the occupation
distribution from $s_1$: $\transp{\1_{s_1}}\sum_t P^t$, where as usual I
omit constants\footnote{To deal with constants, we probably have to be
able to predict \emph{differences} $\psi(s_2)-\psi(s_2')$.}). In
particular, $\E_{s_2\sim \rho}f(s_2)\psi(s_2)$ is supposed to represent
the vector with components $\rho \psi(s_2)$. Then the scalar product in
$E$ represents integration/summation with respect to $\rho$.

Of course we cannot directly learn $\green$ from \eqref{eq:greenrep},
since $\Delta^{-1}$ in the right-hand-side is exactly what we want to
learn. But we can use the Bellman equations \eqref{eq:greenbell}; after
applying \eqref{eq:greenbell} to an arbitrary state $s_1$ and test
function $\psi$, and
substituting the model $\green$ for $\Delta^{-1}$ via
$\eqref{eq:greenrep}$, the left Bellman equation becomes
\begin{gather}
\transp{\1_{s_1}}\Delta^{-1}\psi=\transp{\1_{s_1}}\psi+\transp{\1_{s_1}}P\Delta^{-1}\psi
\qquad \Longrightarrow
\\
\E_{s_2\sim \rho}\,\green(s_1)\cdot
f(s_2)\psi(s_2)
\approx \psi(s_1)+ \E_{s'\sim P(s'|s_1)}\E_{s_2\sim
\rho}\,\green(s')\cdot
f(s_2)\psi(s_2)
\end{gather}
and the right Bellman equation becomes
\begin{gather}
\transp{\1_{s_1}}\Delta^{-1}\psi=\transp{\1_{s_1}}\psi+\transp{\1_{s_1}}\Delta^{-1}P\psi
\qquad \Longrightarrow
\\
\E_{s_2\sim \rho}\,\green(s_1)\cdot
f(s_2)\psi(s_2)
\approx \psi(s_1)+ \E_{s_2\sim \rho}\E_{s'_2\sim P(s'_2|s_2)}
\,\green(s_1)\cdot
f(s_2)\psi(s'_2)
\end{gather}
(indeed, 
$P\psi$ is
represented by $\E_{s_2\sim \rho} f(s_2)(P\psi)(s_2)$, and
$(P\psi)(s_2)=\E_{s'_2\sim P(s'_2|s_2)}\psi(s_2)$).

\emph{Both} these equations are samplable. Therefore, one may
parameterize $f$ and $\green$ by neural networks and train the
parameters. This works off-policy: we have not used that $\rho$ is the
invariant distribution, just some distribution on states we can sample
from.

>>>>

``Universal value function approximators'' (Schaul--Horgan--Gregor--Silver ICML 2015) do a bit of that, but only
with the left (ordinary) Bellman equation. It amounts to solving a number
of
ordinary Bellman equations with different reward functions, and relying on the generalization
abilities of the network, but the right equation is not exploited at all.

@@

If the state space is finite and if we use a one-hot representation, a
solution is indeed $E=\R^{\#S}$ with
\begin{equation}
f(s_2)=\1_{s_2},\qquad
(\green(s_1))_{s_2}=(\Delta^{-1})_{s_1s_2}/\rho(s_2)
\end{equation}
but with infinite or continuous state spaces, we have to find a more
compact representation.

The reason for this form is the following: $E$ is supposed to represent
distributions/functions over states. $f(s_2)$ is supposed to be
mapped to $\1_{s_2}$. $\green(s_1)$ is supposed to be mapped to the
density wrt $\rho$ of the fundamental matrix at $s_1$ (the occupation
distribution from $s_1$: $\transp{\1_{s_1}}\sum_t P^t$, where as usual I
omit constants\footnote{To deal with constants, we probably have to be
able to predict \emph{differences} $\psi(s_2)-\psi(s_2')$, or just use
relative value functions and remove $\psi(s_0)$ for some $s_0$.}). In
particular, $\E_{s_2\sim \rho}f(s_2)\psi(s_2)$ is supposed to represent
the vector with components $\rho \psi(s_2)$. Then the scalar product in
$E$ represents integration/summation with respect to $\rho$.

Of course we cannot directly learn $\green$ from \eqref{eq:greenrep},
since $\Delta^{-1}$ in the right-hand-side is exactly what we want to
learn. But we can use the Bellman equations \eqref{eq:greenbell}; after
applying \eqref{eq:greenbell} to an arbitrary state $s_1$ and test
function $\psi$, and
substituting the model $\green$ for $\Delta^{-1}$ via
$\eqref{eq:greenrep}$, the left Bellman equation becomes
\begin{gather}
\transp{\1_{s_1}}\Delta^{-1}\psi=\transp{\1_{s_1}}\psi+\transp{\1_{s_1}}P\Delta^{-1}\psi
\qquad \Longrightarrow
\\
\E_{s_2\sim \rho}\,\green(s_1)\cdot
f(s_2)\psi(s_2)
\approx \psi(s_1)+ \E_{s'\sim P(s'|s_1)}\E_{s_2\sim
\rho}\,\green(s')\cdot
f(s_2)\psi(s_2)
\end{gather}
because $\transp{\1_{s_1}}P=\E_{s'\sim P(s'|s_1)}\transp{\1_{s'}}\,$.
Meanwhile,
the right Bellman equation becomes
\begin{gather}
\transp{\1_{s_1}}\Delta^{-1}\psi=\transp{\1_{s_1}}\psi+\transp{\1_{s_1}}\Delta^{-1}P\psi
\qquad \Longrightarrow
\\
\E_{s_2\sim \rho}\,\green(s_1)\cdot
f(s_2)\psi(s_2)
\approx \psi(s_1)+ \E_{s_2\sim \rho}\E_{s'_2\sim P(s'_2|s_2)}
\,\green(s_1)\cdot
f(s_2)\psi(s'_2)
\end{gather}
(indeed, 
$P\psi$ is
represented by $\E_{s_2\sim \rho} f(s_2)(P\psi)(s_2)$, and
$(P\psi)(s_2)=\E_{s'_2\sim P(s'_2|s_2)}\psi(s_2)$).

\emph{Both} these equations are samplable. Therefore, one may
parameterize $f$ and $\green$ by neural networks and train the
parameters. This works off-policy: we have not used that $\rho$ is the
invariant distribution, just some distribution on states we can sample
from. Of course, we should apply this to relevant test functions $\psi$,
such as the reward function or any useful (predictive) features, namely,
any function we're interested in predicting the future values of. 
If just using a
random function $\psi$, there is very little signal to train
from. \footnote{An extreme case is to take a purely random $\psi$ at each
training step; then $\psi(s_1)$ and $\psi(s_2)$ are independent unless $s_1=s_2$,
which is a very unlikely event. This amounts to training with white noise
features. Formally it works thanks to the rare case $s_1=s_2$, but
variance is just terrible.}
